---
title: "Tariff Rate Tracker"
author: "The Budget Lab at Yale"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    fig_width: 10
    fig_height: 6
    number_sections: true
  pdf_document:
    toc: true
    number_sections: true
subtitle: "Statutory U.S. Tariff Rates, 2025--2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center'
)

library(tidyverse)
library(here)
library(scales)
library(knitr)
source(here('src', 'helpers.R'))

# Load data
daily_overall   <- read_csv(here('output/daily/daily_overall.csv'))
daily_country   <- read_csv(here('output/daily/daily_by_country.csv'))
daily_authority <- read_csv(here('output/daily/daily_by_authority.csv'))
revision_dates  <- read_csv(here('config/revision_dates.csv'))

# Load country mapping
partner_map <- read_csv(here('resources/country_partner_mapping.csv'))
census_codes <- read_csv(here('resources/census_codes.csv'))

# Load ETR comparison data if available
etr_overall   <- tryCatch(read_csv(here('output/etr/etr_overall.csv')), error = function(e) NULL)
etr_partner   <- tryCatch(read_csv(here('output/etr/etr_by_partner.csv')), error = function(e) NULL)
etr_authority <- tryCatch(read_csv(here('output/etr/etr_by_authority.csv')), error = function(e) NULL)

# Load ETRs comparison data if available
etrs_comp <- tryCatch({
  read_csv(here('output', 'validation', 'etrs_comparison_by_country.csv'),
           col_types = cols(.default = col_character())) %>%
    mutate(
      imports = as.numeric(imports),
      etr = as.numeric(etr),
      etrs_level = as.numeric(etrs_level),
      diff_pp = as.numeric(diff_pp),
      import_share = as.numeric(import_share),
      date = as.Date(date)
    )
}, error = function(e) NULL)

# Policy params
pp <- load_policy_params()
eu27 <- pp$EU27_CODES

# Partner groupings
partner_codes <- c('China' = '5700', 'Canada' = '1220', 'Mexico' = '2010',
                   'Japan' = '5880', 'UK' = '4120')
```

```{r theme}
# Consistent plot theme
theme_tracker <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title = element_text(face = 'bold', size = rel(1.2)),
      plot.subtitle = element_text(color = 'grey40', size = rel(0.9)),
      panel.grid.minor = element_blank(),
      legend.position = 'bottom',
      legend.title = element_blank(),
      plot.caption = element_text(color = 'grey50', size = rel(0.7), hjust = 0)
    )
}

# Color palettes
authority_colors <- c(
  'Section 232'  = '#1b9e77',
  'Section 301'  = '#d95f02',
  'IEEPA'        = '#7570b3',
  'Fentanyl'     = '#e7298a',
  'Section 122'  = '#66a61e',
  'Section 201'  = '#e6ab02'
)

partner_colors <- c(
  'China'   = '#e41a1c',
  'Canada'  = '#377eb8',
  'Mexico'  = '#4daf4a',
  'EU'      = '#984ea3',
  'Japan'   = '#ff7f00',
  'UK'      = '#a65628',
  'FTA ROW' = '#999999',
  'ROW'     = '#666666'
)

partner_labels <- c(
  'china' = 'China', 'canada' = 'Canada', 'mexico' = 'Mexico',
  'eu' = 'EU', 'japan' = 'Japan', 'uk' = 'UK',
  'ftrow' = 'FTA ROW', 'row' = 'ROW'
)
```

# Purpose

The Tariff Rate Tracker constructs statutory U.S. tariff rates at the
HTS-10 product by country level for the 2025--2026 tariff regime. It
processes Harmonized Tariff Schedule (HTS) archives published by the U.S.
International Trade Commission (USITC), along with supplementary policy
resources, to produce a comprehensive panel of tariff rates across all
products and trading partners.

The tracker is designed to support policy analysis by The Budget Lab at
Yale. Its outputs feed directly into The Budget Lab's
[Tariff-Model](https://github.com/Budget-Lab-Yale/Tariff-Model), which
estimates the revenue, distributional, and macroeconomic effects of tariff
policy changes.

The core outputs are:

- An interval-encoded tariff panel covering `r n_distinct(daily_overall$revision)` HTS
  revisions from `r format(min(as.Date(daily_overall$date)), '%B %Y')` through
  `r format(max(as.Date(daily_overall$date)), '%B %Y')`.
- Daily aggregate series decomposed by tariff authority and trading partner.
- Import-weighted effective tariff rate (ETR) estimates at key policy dates.

# Data and Process

## Data sources

The tracker draws on several categories of inputs:

1. **HTS JSON archives** from the USITC. Each revision encodes the full
   tariff schedule: base MFN rates, Chapter 99 additional duties,
   product footnote references, and special program eligibility codes.
   The current build processes
   `r nrow(revision_dates)` revisions.

2. **USITC revision metadata** via the REST API
   (`hts.usitc.gov/reststop/releaseList`). The tracker automatically
   discovers new HTS revisions and appends them to the revision schedule.
   New entries are flagged for manual review because the API returns
   publication dates, which typically lag the policy effective date by
   days to weeks. A SHA-256 hash check on the Chapter 99 PDF provides an
   additional change-detection signal between formal revisions.

3. **Committed resource files** maintained in the repository. These
   include product lists for Section 301 (China), Section 232 derivatives
   and copper, IEEPA-exempt products, fentanyl carve-outs, floor-country
   exemptions, and USMCA product-level utilization shares from USITC
   DataWeb.

4. **Policy parameters** in a single YAML configuration file, encoding
   tariff authority classifications, country-specific rate structures,
   stacking rules, and date-bounded overrides (e.g., Section 122 expiry,
   the Swiss framework agreement).

5. **Census import data** (2024) for computing import-weighted ETRs.
   These weights are optional---the core build produces unweighted
   aggregates without them.

## Processing pipeline

The pipeline runs in seven stages for each HTS revision:

1. **Parse Chapter 99 entries**: extract additional-duty rates, classify
   each entry by tariff authority (Section 232, 301, IEEPA reciprocal,
   fentanyl, Section 122, Section 201).

2. **Parse product master records**: extract ~19,800 HTS-10 product
   lines with base MFN rates, footnote references, and special program
   fields.

3. **Extract policy parameters**: parse Chapter 99 data to determine
   country-specific rate structures, Section 232 rates (including
   blanket vs. country-deal separation for autos), IEEPA phase
   structures, and floor-country applicability.

4. **Build product-country rates**: link products to duties via three
   mechanisms---footnote references (product-specific Ch99 provisions),
   blanket chapter/heading coverage (Section 232), and product list
   matching (Section 301, IEEPA exemptions).

5. **Apply stacking rules**: compute the total tariff for each
   product-country pair. Section 232 takes precedence over IEEPA
   reciprocal on the metal-covered portion; other authorities stack
   according to documented mutual-exclusion rules.

6. **Apply USMCA and exemptions**: reduce rates for Canadian and Mexican
   products proportional to their product-level USMCA utilization share.
   Apply MFN preference adjustments, floor-country product exemptions,
   IEEPA-exempt product lists, and the floor recomputation (which uses
   `ieepa_type` flags to ensure only floor-type entries are adjusted
   against the post-MFN effective base rate).

7. **Aggregate to daily series**: broadcast the piecewise-constant
   revision intervals across calendar days, splitting at policy expiry
   dates (e.g., Section 122's 150-day statutory limit). Both
   exposed-pairs (tariffed only) and all-pairs (full Cartesian panel)
   denominators are computed.

```{r policy-timeline}
# Key policy events for annotation
key_events <- revision_dates %>%
  filter(policy_event != '' & !is.na(policy_event)) %>%
  mutate(date = as.Date(effective_date)) %>%
  filter(date >= '2025-01-01') %>%
  # Select the most important events for annotation
  filter(revision %in% c('basic', 'rev_3', 'rev_6', 'rev_7', 'rev_9',
                          'rev_12', 'rev_17', 'rev_18', 'rev_23',
                          'rev_32', '2026_basic', '2026_rev_4'))
```

# Daily Tariff Rates

## Overall effective tariff rate

```{r fig-overall, fig.height=5}
daily_plot <- daily_overall %>%
  mutate(date = as.Date(date))

# Annotation dates
annotations <- tribble(
  ~date,         ~label,
  '2025-02-04',  'Fentanyl\n(CA/MX/CN)',
  '2025-04-02',  'Liberation\nDay',
  '2025-04-09',  'Phase 1\nsuspended',
  '2025-07-01',  '232\nincrease',
  '2025-08-07',  'Phase 2',
  '2025-11-15',  'Floor\ncountries',
  '2026-02-24',  'IEEPA\ninvalidated'
) %>%
  mutate(date = as.Date(date)) %>%
  left_join(daily_plot %>% distinct(date, .keep_all = TRUE), by = 'date')

has_weights <- !all(is.na(daily_plot$weighted_etr))

p <- ggplot(daily_plot, aes(x = date))

if (has_weights) {
  p <- p + geom_line(aes(y = weighted_etr * 100), color = '#2c5f8a', linewidth = 0.9)
} else {
  p <- p + geom_line(aes(y = mean_total_all_pairs * 100), color = '#2c5f8a', linewidth = 0.9)
}

# Add TPC comparison points if available
if (!is.null(etr_overall) && has_weights) {
  tpc_points <- etr_overall %>%
    filter(!is.na(etr_tpc)) %>%
    mutate(date = as.Date(date))
  p <- p +
    geom_point(data = tpc_points, aes(x = date, y = etr_tpc * 100),
               color = '#d63333', size = 3, shape = 16)
}

yvar <- if (has_weights) 'weighted_etr' else 'mean_total_all_pairs'

p <- p +
  geom_vline(
    data = annotations,
    aes(xintercept = date),
    linetype = 'dashed', color = 'grey60', linewidth = 0.3
  ) +
  geom_text(
    data = annotations,
    aes(x = date, y = Inf, label = label),
    vjust = 1.2, hjust = 0.5, size = 2.5, color = 'grey40', lineheight = 0.85
  ) +
  scale_x_date(date_breaks = '2 months', date_labels = '%b\n%Y') +
  scale_y_continuous(
    limits = c(0, NA),
    labels = function(x) paste0(x, '%')
  ) +
  labs(
    title = 'U.S. Average Tariff Rate (2025--2026)',
    subtitle = if (has_weights) {
      'Import-weighted effective tariff rate; red dots = TPC benchmark'
    } else {
      'Unweighted mean total tariff rate across all product-country pairs'
    },
    x = NULL, y = NULL,
    caption = 'Source: Yale Budget Lab Tariff Rate Tracker. Built from USITC HTS archives.'
  ) +
  theme_tracker()

p
```

The series shows the step-function structure of statutory tariff
changes, with rates only updating when a new HTS revision takes effect.
The fentanyl tariffs on Canada, Mexico, and China in early 2025 are
followed by the April 2 "Liberation Day" reciprocal tariffs, the
subsequent Phase 1 suspension and China de-escalation, and the
reinstatement via Phase 2 in August 2025. The February 2026 drop
reflects the Supreme Court's invalidation of IEEPA tariff authority,
with Section 122 providing a lower replacement rate.

## By tariff authority

```{r fig-authority, fig.height=6}
authority_long <- daily_authority %>%
  mutate(date = as.Date(date)) %>%
  {
    if (has_weights) {
      select(., date, revision,
             `Section 232` = etr_232,
             `Section 301` = etr_301,
             `IEEPA`       = etr_ieepa,
             `Fentanyl`    = etr_fentanyl,
             `Section 122` = etr_s122,
             `Section 201` = etr_section_201)
    } else {
      select(., date, revision,
             `Section 232` = mean_232,
             `Section 301` = mean_301,
             `IEEPA`       = mean_ieepa,
             `Fentanyl`    = mean_fentanyl,
             `Section 122` = mean_s122,
             `Section 201` = mean_section_201)
    }
  } %>%
  pivot_longer(-c(date, revision), names_to = 'authority', values_to = 'rate') %>%
  filter(authority != 'Section 201') # very small, omit for clarity

authority_long$authority <- factor(
  authority_long$authority,
  levels = c('Section 122', 'Fentanyl', 'IEEPA', 'Section 301', 'Section 232')
)

ggplot(authority_long, aes(x = date, y = rate * 100, fill = authority)) +
  geom_area(alpha = 0.85) +
  scale_fill_manual(values = authority_colors, guide = guide_legend(reverse = TRUE)) +
  scale_x_date(date_breaks = '2 months', date_labels = '%b\n%Y') +
  scale_y_continuous(labels = function(x) paste0(x, '%')) +
  labs(
    title = 'U.S. Tariff Rate by Authority (2025--2026)',
    subtitle = if (has_weights) {
      'Import-weighted ETR contribution by tariff authority'
    } else {
      'Unweighted mean rate contribution by tariff authority'
    },
    x = NULL, y = NULL,
    caption = 'Source: Yale Budget Lab Tariff Rate Tracker. Section 201 omitted (< 0.02pp contribution).'
  ) +
  theme_tracker()
```

The authority decomposition illustrates the shifting composition of U.S.
trade policy. Section 232 (steel, aluminum, autos, copper) and Section
301 (China) provide a relatively stable base throughout the period.
IEEPA reciprocal and fentanyl tariffs dominate from April through
February 2026, when they are struck down. Section 122 replaces IEEPA at a
lower overall rate.

## By major trading partner

```{r fig-country-prep}
# Map country codes to names and partner groups
major_partners <- c('5700', '1220', '2010', '5880', '4280', '4120')
major_labels <- c('China', 'Canada', 'Mexico', 'Japan', 'Germany (EU)', 'United Kingdom')
names(major_labels) <- major_partners

country_daily <- daily_country %>%
  mutate(date = as.Date(date)) %>%
  filter(country %in% major_partners) %>%
  mutate(partner = major_labels[as.character(country)])
```

```{r fig-country, fig.height=7}
rate_var <- if (has_weights && !all(is.na(country_daily$weighted_etr))) {
  'weighted_etr'
} else {
  'mean_total_all_pairs'
}

ggplot(country_daily, aes(x = date, y = .data[[rate_var]] * 100, color = partner)) +
  geom_step(linewidth = 0.8) +
  scale_color_manual(values = partner_colors) +
  scale_x_date(date_breaks = '2 months', date_labels = '%b\n%Y') +
  scale_y_continuous(labels = function(x) paste0(x, '%')) +
  labs(
    title = 'Tariff Rate by Major Trading Partner (2025--2026)',
    subtitle = if (rate_var == 'weighted_etr') {
      'Import-weighted effective tariff rate'
    } else {
      'Unweighted mean total tariff rate'
    },
    x = NULL, y = NULL,
    caption = 'Source: Yale Budget Lab Tariff Rate Tracker.'
  ) +
  theme_tracker()
```

```{r fig-country-facet, fig.height=8}
ggplot(country_daily, aes(x = date, y = .data[[rate_var]] * 100)) +
  geom_step(color = '#2c5f8a', linewidth = 0.7) +
  facet_wrap(~ partner, scales = 'free_y', ncol = 2) +
  scale_x_date(date_breaks = '4 months', date_labels = '%b\n%Y') +
  scale_y_continuous(labels = function(x) paste0(x, '%')) +
  labs(
    title = 'Tariff Rate by Partner (Faceted)',
    x = NULL, y = NULL,
    caption = 'Source: Yale Budget Lab Tariff Rate Tracker.'
  ) +
  theme_tracker() +
  theme(strip.text = element_text(face = 'bold'))
```

China faces the highest sustained tariff rates, driven by overlapping
Section 301, IEEPA reciprocal, and fentanyl authorities. Canada and
Mexico are primarily affected by fentanyl tariffs with USMCA providing
partial relief. EU countries, Japan, and the UK see rates jump at
Liberation Day and Phase 2, then fall when IEEPA is invalidated; their
Section 232 exposure on steel, aluminum, and autos persists. After
February 2026, China's rates decline significantly as IEEPA and fentanyl
authorities are vacated, leaving Section 301 and 232 as the remaining
sources of elevated duties.

`r if (!is.null(etr_partner)) '## Partner-level ETR comparison'`

```{r fig-partner-etr, eval=!is.null(etr_partner), fig.height=6}
etr_partner_plot <- etr_partner %>%
  mutate(
    date = as.Date(date),
    partner_label = partner_labels[partner],
    partner_label = factor(partner_label, levels = c(
      'China', 'Canada', 'Mexico', 'EU', 'Japan', 'UK', 'FTA ROW', 'ROW'
    ))
  )

ggplot(etr_partner_plot, aes(x = date, y = etr * 100, color = partner_label)) +
  geom_step(linewidth = 0.8) +
  geom_point(size = 1.5) +
  scale_color_manual(values = partner_colors) +
  scale_x_date(date_breaks = '2 months', date_labels = '%b\n%Y') +
  scale_y_continuous(labels = function(x) paste0(x, '%')) +
  labs(
    title = 'Import-Weighted ETR by Partner Group',
    subtitle = 'At key policy dates',
    x = NULL, y = NULL,
    caption = 'Source: Yale Budget Lab Tariff Rate Tracker, weighted by 2024 Census imports.'
  ) +
  theme_tracker()
```

# Methodology

## Relative to Tax Policy Center (TPC) estimates

The tracker validates its rates against TPC benchmark data at five
snapshot dates corresponding to major policy events. TPC estimates are
derived from the daily rates published by TPC and therefore may reflect
different timing conventions, interpolation methods, or rate assumptions
that are not fully transparent from the published series alone.

```{r tpc-table, eval=!is.null(etr_overall)}
tpc_comparison <- etr_overall %>%
  filter(!is.na(etr_tpc)) %>%
  mutate(
    tracker_pct = sprintf('%.1f%%', etr * 100),
    tpc_pct = sprintf('%.1f%%', etr_tpc * 100),
    diff_pp = sprintf('%+.1f pp', (etr - etr_tpc) * 100)
  ) %>%
  select(Date = date, `Policy Event` = label,
         `Tracker ETR` = tracker_pct,
         `TPC ETR` = tpc_pct,
         Difference = diff_pp)

kable(
  tpc_comparison,
  caption = 'Weighted ETR comparison: Tracker vs. TPC at policy-event dates',
  align = c('l', 'l', 'r', 'r', 'r')
)
```

The main sources of divergence are:

**Section 232/IEEPA mutual exclusion.** The tracker applies mutual
exclusion between Section 232 and IEEPA reciprocal, following the legal
structure where 232 tariffs take precedence. TPC appears to stack IEEPA
reciprocal on top of Section 232 products additively. This accounts for
the largest product-level discrepancies (~25 percentage points on ~920
Chinese steel and aluminum products) and is a deliberate methodological
difference, not a data error.

**Liberation Day timing.** The April 2025 snapshot captures a period of
rapid policy changes (Phase 1 announcement, China escalation, Phase 1
suspension) compressed into a single week. The tracker is tied to HTS
revision dates, which may not align exactly with the dates TPC uses for
its benchmark snapshot. The large divergence at the April 17 date
primarily reflects this timing difference.

**Floor-country treatment.** For countries under a 15% floor rate
structure (EU, Japan, South Korea, Switzerland), the tracker computes
the floor deduction as `max(0, 15% - base_rate)` against the post-MFN
effective base rate (after FTA/GSP exemption shares are applied). An
`ieepa_type` flag preserved from the rate extraction step ensures that
only floor-type entries (not surcharge entries for countries
temporarily outside a floor framework) are recomputed. TPC may use a
different order of operations or exclude duty-free products from floor
treatment.

**USMCA utilization rates.** The tracker uses product-level USMCA
utilization shares, whereas TPC appears to use a different approach
(possibly product-level utilization rates with a different source or
methodology). The tracker's binary-to-continuous shift improved agreement
with TPC for Canadian and Mexican products but differences remain,
particularly for automotive products where rules-of-origin and content
requirements are complex.

```{r tpc-partner-table, eval=!is.null(etr_partner)}
# Latest date with TPC data
tpc_partner_latest <- etr_partner %>%
  filter(!is.na(etr_tpc)) %>%
  filter(date == max(date)) %>%
  mutate(
    partner_label = partner_labels[partner],
    tracker_pct = sprintf('%.1f%%', etr * 100),
    tpc_pct = sprintf('%.1f%%', etr_tpc * 100),
    diff_pp = sprintf('%+.1f pp', (etr - etr_tpc) * 100),
    share_pct = sprintf('%.1f%%', import_share * 100)
  ) %>%
  select(
    Partner = partner_label,
    `Import Share` = share_pct,
    `Tracker ETR` = tracker_pct,
    `TPC ETR` = tpc_pct,
    Difference = diff_pp
  )

kable(
  tpc_partner_latest,
  caption = paste0(
    'Partner-level ETR comparison at latest TPC benchmark date (',
    max(etr_partner$date[!is.na(etr_partner$etr_tpc)]), ')'
  ),
  align = c('l', 'r', 'r', 'r', 'r')
)
```

These comparisons should be interpreted with caution. The TPC benchmark
values used here are derived from TPC's published daily rates at
specific snapshot dates. Any differences may reflect not only
methodological divergence but also differences in timing, trade data
vintages, or product classification that are difficult to disentangle
without access to TPC's underlying product-level data.

`r if (!is.null(etrs_comp)) '## Relative to Tariff-ETRs (Budget Lab prior methodology) {#etrs-comparison}'`

```{r etrs-setup, eval=!is.null(etrs_comp)}
# Dates of interest
focus_dates <- as.Date(c('2026-01-01', '2026-02-24'))
all_dates <- as.Date(c('2026-01-01', '2026-02-24', '2026-07-24'))
```

```{r etrs-overview, eval=!is.null(etrs_comp), results='asis'}
cat('
The Tariff Rate Tracker is a ground-up rebuild of the tariff estimation
infrastructure relative to the **Tariff-ETRs** model (the Budget Lab\'s prior
system). Both models estimate import-weighted average tariff rates using Census
country codes and 2024 HS10-level import data. The tracker introduces the
following changes, which are the primary sources of divergence in the
quantitative comparisons below.

**Rate construction.** The tracker parses rates directly from USITC HTS JSON
archives, with the Chapter 99 parser extracting additional-duty rates, country
scope, and suspension status from the machine-readable tariff schedule.
Tariff-ETRs relied more heavily on manually curated rate inputs and scenario
definitions. New HTS revisions are detected automatically via the USITC REST
API release list; publication-date entries are flagged for manual review until
the correct policy effective date is confirmed.

**232/IEEPA stacking.** The tracker applies mutual exclusion between Section
232 and IEEPA reciprocal, following the legal structure where 232 tariffs take
precedence on the metal-covered portion. Tariff-ETRs stacks additively. This
primarily affects China and is the largest single source of product-level
divergence (~25pp on ~920 steel and aluminum products).

**USMCA utilization.** Tariff-ETRs applied USMCA exemptions using GTAP
sector-level utilization fractions. The tracker uses product-level utilization
shares from USITC DataWeb (programs S/S+), with year-specific files for 2024
and 2025, capturing within-sector variation and the shift from pre-tariff to
tariff-induced utilization patterns. For Section 232 auto and motor vehicle
products, the USMCA share is further scaled by a U.S. content factor (40%).

**MFN preferences.** Tariff-ETRs applies FTA/GSP preference utilization shares
at the product level; the tracker applies a simpler exemption-share adjustment.

**Section 232 derivatives.** The tracker uses BEA Detail Input-Output table
metal-content shares at the HS-10 level, compared to Tariff-ETRs\' simpler
product-level bucketing. This produces more granular scaling of IEEPA rates
on the non-metal portion of derivative aluminum products.

**Section 301 product coverage.** The tracker applies Section 301 as a blanket
tariff using the full USITC \"China Tariffs\" reference product list (~10,400
HTS-8 codes), taking the maximum active rate across overlapping generations.
Tariff-ETRs used exclusive partitioning across 301 lists.

**Section 122 enforcement.** The tracker explicitly models the 150-day
statutory limit, zeroing rates after the expiry date. A configuration flag
allows the expiry to be overridden if Congress extends the authority.

**Section 232 auto deal separation.** The tracker distinguishes blanket auto
tariffs from country-specific deal entries (e.g., bilateral agreements). In
revisions where only deal entries exist, non-deal countries correctly receive
a zero auto rate rather than inheriting the deal rate.

**IEEPA-exempt products.** The tracker includes ITA prefix expansions (~125
additional HTS8 codes in Ch84/85) per US Note 2 subdivision (v)(iii), which
primarily affects Taiwan and Malaysia.
')
```

```{r etrs-overall, eval=!is.null(etrs_comp)}
overall_etrs <- etrs_comp %>%
  group_by(date) %>%
  summarise(
    tracker_rev = sum(etr * imports),
    etrs_rev = sum(etrs_level * imports),
    total_imports = sum(imports),
    .groups = 'drop'
  ) %>%
  mutate(
    `Tracker ETR` = paste0(round(tracker_rev / total_imports * 100, 2), '%'),
    `Tariff-ETRs` = paste0(round(etrs_rev / total_imports * 100, 2), '%'),
    `Difference (pp)` = sprintf('%+.2f', (tracker_rev / total_imports - etrs_rev / total_imports) * 100),
    Date = format(date, '%b %d, %Y')
  ) %>%
  select(Date, `Tracker ETR`, `Tariff-ETRs`, `Difference (pp)`)

kable(overall_etrs, align = 'lrrr', caption = 'Overall import-weighted ETR: Tracker vs. Tariff-ETRs')
```

```{r etrs-overall-text, eval=!is.null(etrs_comp), results='asis'}
cat('
The overall gap ranges from -0.2pp to +0.7pp across the three dates. The
February and July dates are within 0.2pp. January is wider (+0.7pp) partly
because the import match rate is lower (73% vs 98%) due to product
renumbering between the import data vintage and the January HTS revision.
After February 2026, remaining gaps are primarily USMCA and MFN preference
differences (where Tariff-ETRs tends to be higher).
')
```

```{r etrs-helpers, eval=!is.null(etrs_comp)}
make_country_table <- function(comp, d, eu_codes) {
  at_date <- comp %>% filter(date == d)

  # Named partners
  partner_rows <- tibble(code = names(partner_codes), census = unname(partner_codes)) %>%
    left_join(at_date, by = c('census' = 'country')) %>%
    transmute(
      Partner = code,
      `Imports ($B)` = round(imports / 1e9, 1),
      `Tracker ETR` = paste0(round(etr * 100, 2), '%'),
      `Tariff-ETRs` = paste0(round(etrs_level * 100, 2), '%'),
      `Diff (pp)` = sprintf('%+.2f', diff_pp)
    )

  # EU aggregate
  eu_data <- at_date %>% filter(country %in% eu_codes)
  if (nrow(eu_data) > 0) {
    eu_imp <- sum(eu_data$imports)
    eu_tracker <- sum(eu_data$etr * eu_data$imports) / eu_imp
    eu_etrs <- sum(eu_data$etrs_level * eu_data$imports) / eu_imp
    eu_row <- tibble(
      Partner = 'EU-27',
      `Imports ($B)` = round(eu_imp / 1e9, 1),
      `Tracker ETR` = paste0(round(eu_tracker * 100, 2), '%'),
      `Tariff-ETRs` = paste0(round(eu_etrs * 100, 2), '%'),
      `Diff (pp)` = sprintf('%+.2f', (eu_tracker - eu_etrs) * 100)
    )
    partner_rows <- bind_rows(partner_rows, eu_row)
  }

  # Overall
  total_imp <- sum(at_date$imports)
  overall_tracker <- sum(at_date$etr * at_date$imports) / total_imp
  overall_etrs <- sum(at_date$etrs_level * at_date$imports) / total_imp
  overall_row <- tibble(
    Partner = '**Overall**',
    `Imports ($B)` = round(total_imp / 1e9, 1),
    `Tracker ETR` = paste0(round(overall_tracker * 100, 2), '%'),
    `Tariff-ETRs` = paste0(round(overall_etrs * 100, 2), '%'),
    `Diff (pp)` = sprintf('%+.2f', (overall_tracker - overall_etrs) * 100)
  )

  bind_rows(partner_rows, overall_row)
}

make_top_table <- function(comp, d, n = 15) {
  census_names <- tryCatch(
    read_csv(here('resources', 'census_codes.csv'),
             col_types = cols(.default = col_character())),
    error = function(e) NULL
  )

  at_date <- comp %>%
    filter(date == d) %>%
    arrange(desc(etr)) %>%
    head(n)

  if (!is.null(census_names)) {
    at_date <- at_date %>%
      left_join(census_names %>% select(Code, Name), by = c('country' = 'Code'))
  } else {
    at_date <- at_date %>% mutate(Name = country)
  }

  at_date %>%
    transmute(
      Country = coalesce(Name, country),
      `Imports ($B)` = round(imports / 1e9, 1),
      `Tracker ETR` = paste0(round(etr * 100, 1), '%'),
      `Tariff-ETRs` = paste0(round(etrs_level * 100, 1), '%'),
      `Diff (pp)` = sprintf('%+.1f', diff_pp)
    )
}
```

```{r etrs-country-jan-header, eval=!is.null(etrs_comp), results='asis'}
cat('\n### January 1, 2026\n')
```

```{r etrs-country-jan, eval=!is.null(etrs_comp)}
jan_tbl <- make_country_table(etrs_comp, as.Date('2026-01-01'), eu27)
kable(jan_tbl, align = 'lrrrr',
      caption = 'Country-level ETR comparison -- January 1, 2026')
```

```{r etrs-jan-text, eval=!is.null(etrs_comp), results='asis'}
cat('
**Notable gaps (Jan 1):**

- **China (+5.3pp):** Driven by 232/IEEPA stacking difference. Tracker applies
  mutual exclusion on steel/aluminum; Tariff-ETRs stacks additively.
- **Canada (+4.6pp):** USMCA utilization -- Tracker uses binary eligibility
  (~24% of products), Tariff-ETRs uses ~50% median utilization rate.
- **Japan (-2.3pp):** MFN preference shares -- Tariff-ETRs applies FTA preference
  utilization for Japan--US trade.
- **EU (-1.6pp):** MFN preference shares for EU trade.
')
```

```{r etrs-country-feb-header, eval=!is.null(etrs_comp), results='asis'}
cat('\n### February 24, 2026\n')
```

```{r etrs-country-feb, eval=!is.null(etrs_comp)}
feb_tbl <- make_country_table(etrs_comp, as.Date('2026-02-24'), eu27)
kable(feb_tbl, align = 'lrrrr',
      caption = 'Country-level ETR comparison -- February 24, 2026')
```

```{r etrs-feb-text, eval=!is.null(etrs_comp), results='asis'}
cat('
**Notable gaps (Feb 24):**

- **China (+5.5pp):** Stacking difference persists. With IEEPA invalidated, the
  gap is now driven by 232/301 stacking and MFN preference differences rather
  than IEEPA stacking.
- **Canada (-1.6pp):** Sign reversal -- with IEEPA invalidated, the USMCA effect
  is smaller, and MFN preference differences dominate.
- **Japan (-1.6pp), EU (-0.8pp):** MFN preference differences. The floor fix
  (Step 6d `ieepa_type` guard) tightened these gaps from prior estimates.
')
```

```{r etrs-top-header, eval=!is.null(etrs_comp), results='asis'}
cat('\n### Top 15 countries by Tracker ETR\n')
```

```{r etrs-top-jan, eval=!is.null(etrs_comp)}
kable(make_top_table(etrs_comp, as.Date('2026-01-01')),
      caption = 'Top 15 countries by Tracker ETR -- January 1, 2026')
```

```{r etrs-top-feb, eval=!is.null(etrs_comp)}
kable(make_top_table(etrs_comp, as.Date('2026-02-24')),
      caption = 'Top 15 countries by Tracker ETR -- February 24, 2026')
```

```{r etrs-authority-header, eval=!is.null(etrs_comp), results='asis'}
cat('\n### Authority decomposition at benchmark dates\n')
```

```{r etrs-authority-decomp, eval=!is.null(etrs_comp)}
auth_rows <- daily_authority %>%
  mutate(date = as.Date(date)) %>%
  filter(date %in% focus_dates) %>%
  select(date, revision,
         etr_232, etr_301, etr_ieepa, etr_fentanyl,
         etr_s122, etr_section_201, etr_other)

if (nrow(auth_rows) == 0) {
  for (d in focus_dates) {
    nearest <- daily_authority %>%
      mutate(date = as.Date(date), dist = abs(as.numeric(date - d))) %>%
      slice_min(dist, n = 1)
    auth_rows <- bind_rows(auth_rows, nearest %>%
      select(date, revision,
             etr_232, etr_301, etr_ieepa, etr_fentanyl,
             etr_s122, etr_section_201, etr_other))
  }
}

auth_display <- auth_rows %>%
  mutate(
    Date = format(date, '%b %d, %Y'),
    `Section 232` = paste0(round(etr_232 * 100, 2), '%'),
    `Section 301` = paste0(round(etr_301 * 100, 2), '%'),
    `IEEPA Reciprocal` = paste0(round(etr_ieepa * 100, 2), '%'),
    `IEEPA Fentanyl` = paste0(round(etr_fentanyl * 100, 2), '%'),
    `Section 122` = paste0(round(etr_s122 * 100, 2), '%'),
    `Section 201` = paste0(round(etr_section_201 * 100, 2), '%'),
    Total = paste0(round((etr_232 + etr_301 + etr_ieepa + etr_fentanyl +
                          etr_s122 + etr_section_201 + etr_other) * 100, 2), '%')
  ) %>%
  select(Date, `Section 232`, `Section 301`, `IEEPA Reciprocal`,
         `IEEPA Fentanyl`, `Section 122`, `Section 201`, Total)

kable(auth_display, align = 'lrrrrrrr',
      caption = 'Tracker ETR decomposition by tariff authority (import-weighted)')
```

```{r etrs-sources, eval=!is.null(etrs_comp), results='asis'}
cat('
### Data sources

- **Tracker snapshots**: Built from USITC HTS JSON archives via the tariff-rate-tracker
  pipeline (`00_build_timeseries.R`). Rates reflect statutory tariff schedule
  including Chapter 99 additional duties.
- **Tariff-ETRs benchmark**: `2-21_temp` scenario from the Tariff-ETRs model,
  providing country-level ETRs at 3 dates.
- **Import weights**: HS10-by-country 2024 Census trade data (from Tariff-ETRs cache),
  total imports = $3,124B.
- **HTS concordance**: 266 product code mappings across revisions, used to remap
  import weights when codes are renumbered between the import data vintage and
  a given HTS revision.
- **IEEPA invalidation**: Modeled as of February 24, 2026 (SCOTUS ruling). Both
  reciprocal and fentanyl IEEPA rates zeroed on and after that date.
- **Section 122 expiry**: Statutory 150-day limit; effective February 24, expires
  July 23, 2026.
')
```

---

*Generated by `docs/analysis/tariff_rate_tracker_overview.Rmd` on `r Sys.time()`.*
