---
title: "Tariff Rate Tracker vs. Tariff-ETRs Comparison"
author: "Tariff Rate Tracker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    number_sections: true
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(knitr)
source(here('src', 'helpers.R'))
```

# Overview

This document compares the effective tariff rates (ETRs) produced by the
**Tariff Rate Tracker** with those from the **Tariff-ETRs** model at three
benchmark dates. Both models estimate the import-weighted average tariff rate
using Census country codes and 2024 HS10-level import data.

**Key methodological differences:**

- **232/IEEPA stacking**: The Tracker applies mutual exclusion (232 displaces
  IEEPA on metal products); Tariff-ETRs stacks additively. This primarily
  affects China.
- **USMCA**: The Tracker uses binary S/S+ eligibility from HTS; Tariff-ETRs
  uses product-level utilization rates (~50% median for CA/MX).
- **MFN preferences**: Tariff-ETRs applies FTA/GSP preference utilization shares;
  the Tracker applies a simpler exemption-share adjustment.
- **301 product lists**: Minor differences in list coverage.

```{r load-data}
# Load ETRs comparison output
etrs_comp <- read_csv(here('output', 'validation', 'etrs_comparison_by_country.csv'),
                      col_types = cols(.default = col_character())) %>%
  mutate(
    imports = as.numeric(imports),
    etr = as.numeric(etr),
    etrs_level = as.numeric(etrs_level),
    diff_pp = as.numeric(diff_pp),
    import_share = as.numeric(import_share),
    date = as.Date(date)
  )

# Load daily by-authority for weighted ETR decomposition
daily_auth <- read_csv(here('output', 'daily', 'daily_by_authority.csv'),
                       col_types = cols(.default = col_guess()))

# Load daily by-country
daily_cty <- read_csv(here('output', 'daily', 'daily_by_country.csv'),
                      col_types = cols(.default = col_guess()))

# Policy params for labels
pp <- load_policy_params()
eu27 <- pp$EU27_CODES

# Partner groupings
partner_codes <- c('China' = '5700', 'Canada' = '1220', 'Mexico' = '2010',
                   'Japan' = '5880', 'UK' = '4120')

# Dates of interest
focus_dates <- as.Date(c('2026-01-01', '2026-02-24'))
all_dates <- as.Date(c('2026-01-01', '2026-02-24', '2026-07-24'))
```

# Overall ETR Comparison

```{r overall-summary}
overall <- etrs_comp %>%
  group_by(date) %>%
  summarise(
    tracker_rev = sum(etr * imports),
    etrs_rev = sum(etrs_level * imports),
    total_imports = sum(imports),
    .groups = 'drop'
  ) %>%
  mutate(
    `Tracker ETR` = paste0(round(tracker_rev / total_imports * 100, 2), '%'),
    `Tariff-ETRs` = paste0(round(etrs_rev / total_imports * 100, 2), '%'),
    `Difference (pp)` = sprintf('%+.2f', (tracker_rev / total_imports - etrs_rev / total_imports) * 100),
    Date = format(date, '%b %d, %Y')
  ) %>%
  select(Date, `Tracker ETR`, `Tariff-ETRs`, `Difference (pp)`)

kable(overall, align = 'lrrr', caption = 'Overall import-weighted ETR comparison')
```

The overall gap ranges from -0.5pp to +0.6pp across the three dates. The sign
flip at July 24 reflects the post-IEEPA-invalidation and post-Section 122
expiry environment, where remaining gaps are primarily USMCA and MFN preference
differences (where Tariff-ETRs tends to be higher).

# Country-Level Comparison

## January 1, 2026

```{r country-jan}
make_country_table <- function(comp, d, eu_codes) {
  at_date <- comp %>% filter(date == d)

  # Named partners
  partner_rows <- tibble(code = names(partner_codes), census = unname(partner_codes)) %>%
    left_join(at_date, by = c('census' = 'country')) %>%
    transmute(
      Partner = code,
      `Imports ($B)` = round(imports / 1e9, 1),
      `Tracker ETR` = paste0(round(etr * 100, 2), '%'),
      `Tariff-ETRs` = paste0(round(etrs_level * 100, 2), '%'),
      `Diff (pp)` = sprintf('%+.2f', diff_pp)
    )

  # EU aggregate
  eu_data <- at_date %>% filter(country %in% eu_codes)
  if (nrow(eu_data) > 0) {
    eu_imp <- sum(eu_data$imports)
    eu_tracker <- sum(eu_data$etr * eu_data$imports) / eu_imp
    eu_etrs <- sum(eu_data$etrs_level * eu_data$imports) / eu_imp
    eu_row <- tibble(
      Partner = 'EU-27',
      `Imports ($B)` = round(eu_imp / 1e9, 1),
      `Tracker ETR` = paste0(round(eu_tracker * 100, 2), '%'),
      `Tariff-ETRs` = paste0(round(eu_etrs * 100, 2), '%'),
      `Diff (pp)` = sprintf('%+.2f', (eu_tracker - eu_etrs) * 100)
    )
    partner_rows <- bind_rows(partner_rows, eu_row)
  }

  # Overall
  total_imp <- sum(at_date$imports)
  overall_tracker <- sum(at_date$etr * at_date$imports) / total_imp
  overall_etrs <- sum(at_date$etrs_level * at_date$imports) / total_imp
  overall_row <- tibble(
    Partner = '**Overall**',
    `Imports ($B)` = round(total_imp / 1e9, 1),
    `Tracker ETR` = paste0(round(overall_tracker * 100, 2), '%'),
    `Tariff-ETRs` = paste0(round(overall_etrs * 100, 2), '%'),
    `Diff (pp)` = sprintf('%+.2f', (overall_tracker - overall_etrs) * 100)
  )

  bind_rows(partner_rows, overall_row)
}

jan_tbl <- make_country_table(etrs_comp, as.Date('2026-01-01'), eu27)
kable(jan_tbl, align = 'lrrrr',
      caption = 'Country-level ETR comparison — January 1, 2026')
```

**Notable gaps (Jan 1):**

- **China (+5.2pp):** Driven by 232/IEEPA stacking difference. Tracker applies
  mutual exclusion on steel/aluminum; Tariff-ETRs stacks additively.
- **Canada (+4.5pp):** USMCA utilization — Tracker uses binary eligibility
  (~24% of products), Tariff-ETRs uses ~50% median utilization rate.
- **Japan (-2.3pp):** MFN preference shares — Tariff-ETRs applies FTA preference
  utilization for Japan–US trade.
- **EU (-1.5pp):** MFN preference shares for EU trade.

## February 24, 2026

```{r country-feb}
feb_tbl <- make_country_table(etrs_comp, as.Date('2026-02-24'), eu27)
kable(feb_tbl, align = 'lrrrr',
      caption = 'Country-level ETR comparison — February 24, 2026')
```

**Notable gaps (Feb 24):**

- **China (+4.4pp):** Similar stacking difference, narrower because IEEPA
  reciprocal is invalidated on this date (SCOTUS ruling), reducing the base
  that the stacking difference applies to.
- **Canada (-1.5pp):** Sign reversal — with IEEPA invalidated, the USMCA effect
  is smaller, and MFN preference differences dominate.
- **Japan (+1.8pp), UK (+1.9pp):** Section 122 treatment differences.

# Authority Decomposition

The Tracker's daily series decomposes the ETR into contributions by tariff
authority. The table below shows the import-weighted ETR contribution from
each authority at the two focus dates.

```{r authority-decomp}
# Find the closest available dates in daily_auth
auth_rows <- daily_auth %>%
  filter(date %in% focus_dates) %>%
  select(date, revision,
         etr_232, etr_301, etr_ieepa, etr_fentanyl,
         etr_s122, etr_section_201, etr_other)

if (nrow(auth_rows) == 0) {
  # Try nearest dates
  for (d in focus_dates) {
    nearest <- daily_auth %>%
      mutate(dist = abs(as.numeric(date - d))) %>%
      slice_min(dist, n = 1)
    auth_rows <- bind_rows(auth_rows, nearest %>%
      select(date, revision,
             etr_232, etr_301, etr_ieepa, etr_fentanyl,
             etr_s122, etr_section_201, etr_other))
  }
}

auth_display <- auth_rows %>%
  mutate(
    Date = format(date, '%b %d, %Y'),
    `Section 232` = paste0(round(etr_232 * 100, 2), '%'),
    `Section 301` = paste0(round(etr_301 * 100, 2), '%'),
    `IEEPA Reciprocal` = paste0(round(etr_ieepa * 100, 2), '%'),
    `IEEPA Fentanyl` = paste0(round(etr_fentanyl * 100, 2), '%'),
    `Section 122` = paste0(round(etr_s122 * 100, 2), '%'),
    `Section 201` = paste0(round(etr_section_201 * 100, 2), '%'),
    Total = paste0(round((etr_232 + etr_301 + etr_ieepa + etr_fentanyl +
                          etr_s122 + etr_section_201 + etr_other) * 100, 2), '%')
  ) %>%
  select(Date, `Section 232`, `Section 301`, `IEEPA Reciprocal`,
         `IEEPA Fentanyl`, `Section 122`, `Section 201`, Total)

kable(auth_display, align = 'lrrrrrrr',
      caption = 'Tracker ETR decomposition by tariff authority (import-weighted)')
```

# Top Country ETRs

```{r top-countries}
make_top_table <- function(comp, d, n = 15) {
  # Load census names
  census <- tryCatch(
    read_csv(here('resources', 'census_codes.csv'),
             col_types = cols(.default = col_character())),
    error = function(e) NULL
  )

  at_date <- comp %>%
    filter(date == d) %>%
    arrange(desc(etr)) %>%
    head(n)

  if (!is.null(census)) {
    at_date <- at_date %>%
      left_join(census %>% select(Code, Name), by = c('country' = 'Code'))
  } else {
    at_date <- at_date %>% mutate(Name = country)
  }

  at_date %>%
    transmute(
      Country = coalesce(Name, country),
      `Imports ($B)` = round(imports / 1e9, 1),
      `Tracker ETR` = paste0(round(etr * 100, 1), '%'),
      `Tariff-ETRs` = paste0(round(etrs_level * 100, 1), '%'),
      `Diff (pp)` = sprintf('%+.1f', diff_pp)
    )
}
```

## January 1, 2026 — Top 15 by Tracker ETR

```{r top-jan}
kable(make_top_table(etrs_comp, as.Date('2026-01-01')),
      caption = 'Top 15 countries by Tracker ETR — January 1, 2026')
```

## February 24, 2026 — Top 15 by Tracker ETR

```{r top-feb}
kable(make_top_table(etrs_comp, as.Date('2026-02-24')),
      caption = 'Top 15 countries by Tracker ETR — February 24, 2026')
```

# Data Sources and Methodology Notes

- **Tracker snapshots**: Built from USITC HTS JSON archives via the tariff-rate-tracker
  pipeline (`00_build_timeseries.R`). Rates reflect statutory tariff schedule
  including Chapter 99 additional duties.
- **Tariff-ETRs benchmark**: `2-21_temp` scenario from the Tariff-ETRs model,
  providing country-level ETRs at 3 dates.
- **Import weights**: HS10-by-country 2024 Census trade data (from Tariff-ETRs cache),
  total imports = $3,124B.
- **IEEPA invalidation**: Modeled as of February 24, 2026 (SCOTUS ruling). Both
  reciprocal and fentanyl IEEPA rates zeroed on and after that date.
- **Section 122 expiry**: Statutory 150-day limit; effective February 24, expires
  July 23, 2026.

---

*Generated by `docs/etrs_comparison_report.Rmd` on `r Sys.time()`.*
*Comparison data: `output/validation/etrs_comparison_by_country.csv`.*
